---
title: "R Notebook"
author: "Zongyu Wu"
date: "Spring 2024"
output:
  pdf_document: default
  html_document:
    df_print: paged
email: zongyuwu97@gmail.com
subtitle: Practicum II CS5200
---

## Connect to database

```{r connectDatabase}
# 1. Library
library("RMariaDB")
library("DBI")
library(ggplot2)

# 2. Settings
db_user <- 'admin'            # use your value from the setup
db_password <- 'wuzongyu'    # use your value from the setup
db_name <- 'practicum1'         # use your value from the setup

db_host <- 'cs5200.ceko8sdfannu.us-east-1.rds.amazonaws.com'       # for aws

db_port <- 3306

# 3. Connect to remote server database

db.aws <-  DBI::dbConnect(RMariaDB::MariaDB(), user = db_user, password = db_password,
                           dbname = db_name, host = db_host, port = db_port)
# 4. Knit setting
knitr::opts_chunk$set(connection = "db.aws", echo = FALSE, warning = FALSE, message = FALSE)
```

## Analytical Query I: Top five sales reps with the most sales broken down by year. So, for each year for which there is data, the top five reps for that year. Think about how you would best present this information so it is easily usable by an analyst or manager.

```{r}
years <- dbGetQuery(db.aws, "select distinct year from rep_facts")

for (year in years) {
  sql <- "
    select rep_name, sum(total_amount_sold) as total_amount_sold, year
    from rep_facts
    where year = ?
    group by rep_name
    order by year asc, sum(total_amount_sold) desc
    limit 5
  "
  res <- dbGetQuery(db.aws, sql, list(year))
  print(res)
}
```

Above shows the total amount sold for the top 5 sales.

## Analytical Query II: Total sold per product per quarter. Think about how to best display this information.

```{sql connection=db.aws}
select product_name, quarter, sum(total_amount_sold) as total_amount_sold
from product_facts
group by product_name, quarter
```


## Analytical Query III: Number of units sold per product per region. Show the result in a line graph visualization.

```{r}
sql <- "
select product_name, region, sum(total_unit) as total_unit
from product_facts
group by product_name, region
"

res <- dbGetQuery(db.aws, sql)
res$region <- factor(res$region, levels = unique(res$region))

# Create the line chart with only points
# Create the line chart with lines connecting points for each product_name
ggplot(res, aes(x = region, y = total_unit, color = product_name, group = product_name)) +
  geom_line() +  # Connect points with lines
  geom_point() +  # Show individual points
  labs(title = "Total Unit by Region for Each Product",
       x = "Region",
       y = "Total Unit",
       color = "Product Name") +
  theme_minimal()
```

## Analytical Query IV: Average sales per sales rep over the years. Show the result in a line graph visualization.

```{r}
sql <- "
with total_year as (
select sum(total_amount_sold) as total_amount_sold, rep_name, year
from rep_facts
group by year, rep_name
)
SELECT
    rep_name,
    AVG(total_amount_sold) AS avg_sales
FROM
    total_year
GROUP BY
    rep_name
"
res <- dbGetQuery(db.aws, sql)

# Create the line chart
ggplot(res, aes(x = rep_name, y = avg_sales, group = 1)) +
  geom_line() +
  geom_point() +
  labs(title = "Average Sales per Sales Rep Over the Years",
       x = "Sales Rep",
       y = "Average Sales") +
  theme_minimal()
```



## Clean up

```{r cleanup}
status <- dbDisconnect(mydb.aws)
```