---
title: "WuZ.CS5200.Report.Sp24.sqldf"
author: "Zongyu Wu"
date: "Spring 2024"
output:
  html_document:
    df_print: paged
---

## Preparation
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```
Load data from `.csv` files and bind them into a single data frame for further analysis.

```{r}
library(sqldf)
s2o <- read.csv("https://s3.us-east-2.amazonaws.com/artificium.us/assignments/80.xml/a-80-305/gen-xml/synthsalestxns-Sep2Oct.csv")
n2d <- read.csv("https://s3.us-east-2.amazonaws.com/artificium.us/assignments/80.xml/a-80-305/gen-xml/synthsalestxns-Nov2Dec.csv")
j2m <- read.csv("https://s3.us-east-2.amazonaws.com/artificium.us/assignments/80.xml/a-80-305/gen-xml/synthsalestxns-Jan2Mar.csv")

df <- rbind(s2o, rbind(n2d, j2m))
df$date <- strftime(as.Date(df$date, "%m/%d/%y"), "%y-%m-%d")

head(df, 5)
```

## Calculation

Calculate data needed in the analysis.
Get number of visits

```{r}
# Number of visits
NumberOfVisits <- nrow(df)
print(NumberOfVisits)
```
Get number of restaurants
```{r}
# Number of restaurants
sql <- "select count(distinct restaurant) as res from df"
NumberOfRestaurants <- sqldf(sql)
print(NumberOfRestaurants)
```
Get all dates
```{r}
# All dates
sql <- "select date from df 
where date is not null 
order by date asc"
dates <- sqldf(sql)
head(dates, 5)
```
Get average spent per visit across restaurants
```{r}
# Average spent per visit across restaurants
sql <- "select sum(cast(substring(amount, 2) as decimal(10, 2))) / count(*) as avg
from df 
where amount is not null"
avg <- sqldf(sql)
head(avg)
```
Get standard deviation of all spent
```{r}
# Standard deviation
sql <- "select stdev(cast(substring(amount, 2) as decimal(10, 2))) as stdev
from df 
where amount is not null"
stdev <- sqldf(sql)
head(stdev)
```
Get most visited restaurant
```{r}
# Most visited restaurant
sql <- "select count(*) as visits, restaurant
from df 
group by restaurant
order by visits desc"
visits <- sqldf(sql)
head(visits)
```
Get revenue by restaurant
```{r}
# Revenue by year
sql <- "select sum(cast(substring(amount, 2) as decimal(10, 2))) as revenue, restaurant
from df 
group by restaurant
order by revenue desc"
revenue <- sqldf(sql)
head(revenue)
```
Get average size of a party 
```{r}
# Average size of a party 
sql <- "select sum(guests) * 1.0 / count(*)  as partySize
from df 
where guests is not null"
size <- sqldf(sql)
head(size)
```
Get percentage of party consumed wine
```{r}
# Percent of party consumed wine
sql <- "select (count(case when wine = \"Yes\" then 1 end) * 100.0) / count(wine) as percent
from df
where wine is not null"
wine <- sqldf(sql)
head(wine)
```
Get summary of visits and revenue for each restaurant
```{r}
# Summary of visits and revenue for each restaurant
sql <- "select restaurant, count(*) as visits, sum(cast(substring(amount, 2) as decimal(10, 2))) * 1.0 / count(*) as avgRevenue
from df
where amount is not null
group by restaurant"
summary <- sqldf(sql)
head(summary)
```

## Analysis of Restaurant Visits

We recorded a total number of `r NumberOfVisits` visits across `r NumberOfRestaurants[1, 1]` different restaurants from `r dates[1, 1]` to `r dates[nrow(df), 1]`. The average amount spent for a visit across all restaurants is `r avg[1, 1]` (σ = `r stdev[1, 1]`). The most visits were at `r visits[1, 2]` while the largest revenue across all years was at `r revenue[1, 2]`. The average size of a party was `r size[1, 1]` and `r wine[1, 1]`% of all parties consumed wine.

### Summary of Revenue
```{r}
knitr::kable(summary)
```
