---
title: "Assignment / Explore Query Planning and Indexing"
author: "Zongyu WU"
date: "Spring 2024"
output:
  pdf_document: default
  html_notebook: default
---

Setup connection.
```{r}
library(RSQLite)
dbcon <- dbConnect(RSQLite::SQLite(), "sakila.db")
```

## Question 1
```{r}
# Drop index
dbExecute(dbcon, "drop index if exists TitleIndex")
```


```{r}
query <- "select l.NAME as LanguageName, count(f.FILM_ID) as NumberOfFilms
  from LANGUAGE l join FILM f on l.LANGUAGE_ID = f.LANGUAGE_ID
  group by l.LANGUAGE_ID"
res <- dbGetQuery(dbcon, query)
print(res)
```

## Question 2
```{r}
query <- "EXPLAIN QUERY PLAN
  select l.NAME as LanguageName, count(f.FILM_ID) as NumberOfFilms
  from LANGUAGE l join FILM f on l.LANGUAGE_ID = f.LANGUAGE_ID
  group by l.LANGUAGE_ID"
res <- dbGetQuery(dbcon, query)
print(res)
```

## Question 3
```{r}
before3 <- Sys.time()
query <- "select f.TITLE, c.NAME, f.LENGTH
  from FILM f join FILM_CATEGORY fc on f.FILM_ID = fc.FILM_ID
  join CATEGORY c on c.CATEGORY_ID = fc.CATEGORY_ID
  where TITLE =  \"ZORRO ARK\""
res <- dbGetQuery(dbcon, query)
after3 <- Sys.time()
print(res)
```
## Question 4
```{r}
query <- "EXPLAIN QUERY PLAN
  select f.TITLE, c.NAME, f.LENGTH
  from FILM f join FILM_CATEGORY fc on f.FILM_ID = fc.FILM_ID
  join CATEGORY c on c.CATEGORY_ID = fc.CATEGORY_ID
  where TITLE =  \"ZORRO ARK\""
res <- dbGetQuery(dbcon, query)
print(res)
```

## Question 5
```{r}
query <- "create index if not exists TitleIndex on FILM(TITLE)"
res <- dbExecute(dbcon, query)
print(res)
```

## Question 6
```{r}
before6 <- Sys.time()
query <- "select f.TITLE, c.NAME, f.LENGTH
  from FILM f join FILM_CATEGORY fc on f.FILM_ID = fc.FILM_ID
  join CATEGORY c on c.CATEGORY_ID = fc.CATEGORY_ID
  where TITLE =  \"ZORRO ARK\""
res <- dbGetQuery(dbcon, query)
after6 <- Sys.time()
print(res)
```

```{r}
query <- "EXPLAIN QUERY PLAN
  select f.TITLE, c.NAME, f.LENGTH
  from FILM f join FILM_CATEGORY fc on f.FILM_ID = fc.FILM_ID
  join CATEGORY c on c.CATEGORY_ID = fc.CATEGORY_ID
  where TITLE =  \"ZORRO ARK\""
res <- dbGetQuery(dbcon, query)
print(res)
```

## Question 7
It's not the same. After adding the index, the search on f is changed from using integer primary key to using index TitleIndex. The scan on fc using covering index is also changed to a search using covering index. The query plan will show in the detail that it used the index. By adding an index, related plans are changed from scan to search. And search on that specific column is using that index for searching.


## Question 8
I added the time calculation function in Question 3 and Question 6 to save the code repeat. Here I will print the results from the above two questions.
```{r}
cat("Question 3 time elapsed: ", round((after3 - before3) ,3), " sec\n")
cat("Question 6 time elapsed: ", round((after6 - before6) ,3), " sec")
```
As we can see, the time is reduced. The amount of reduce varies though. It may be the same query is parsed in different ways.

## Question 9
```{r}
query <- "select f.TITLE, l.NAME, f.LENGTH
  from FILM f join LANGUAGE l on f.LANGUAGE_ID = l.LANGUAGE_ID
  where lower(f.TITLE) like '%gold%'"
res <- dbGetQuery(dbcon, query)
print(res)
```

## Question 10
```{r}
query <- "EXPLAIN QUERY PLAN
  select f.TITLE, l.NAME, f.LENGTH
  from FILM f join LANGUAGE l on f.LANGUAGE_ID = l.LANGUAGE_ID
  where lower(f.TITLE) like '%gold%'"
res <- dbGetQuery(dbcon, query)
print(res)
```
It doesn't use the index. It's because pattern match LIKE is used here.

## Clean
```{r}
dbDisconnect(dbcon)
```

